\documentclass{article}

\usepackage{graphicx} % Required for inserting images
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath}
\usepackage{amsthm} %proof environment
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{enumitem} %nice lists
\usepackage{verbatim} %useful for something 
\usepackage{xcolor}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{blindtext} % I have no idea what this is 
\usepackage{caption}  % need this for unnumbered captions/figures
\usepackage{natbib}
\usepackage{tikz}
\usepackage{hyperref}

\titleformat{\section}
{\bfseries\Large}
{Lecture \thesection:}
{5pt}{}

\begin{document}

\title{AM 275 - Magnetohydrodynamics: Lecture Notes}
\author{Dante Buhl}
\doublespacing


\newcommand{\wrms}{w_{\text{rms}}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\tb}[1]{\textbf{#1}}
\newcommand{\bmp}[1]{\begin{minipage}{#1\textwidth}}
\newcommand{\emp}{\end{minipage}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\K}{\bs{\mathrm{K}}}
\newcommand{\m}{\bs{\mu}_*}
\newcommand{\s}{\bs{\Sigma}_*}
\newcommand{\dt}{\Delta t}
\newcommand{\dx}{\Delta x}
\newcommand{\tr}[1]{\text{Tr}(#1)}
\newcommand{\Tr}[1]{\text{Tr}(#1)}
\newcommand{\Div}{\nabla \cdot}
\renewcommand{\div}{\nabla \cdot}
\newcommand{\Curl}{\nabla \times}
\newcommand{\Grad}{\nabla}
\newcommand{\grad}{\nabla}
\newcommand{\grads}{\nabla_s}
\newcommand{\gradf}{\nabla_f}
\newcommand{\xs}{x_s}
\newcommand{\xf}{x_f}
\newcommand{\ts}{t_s}
\newcommand{\tf}{t_f}
\newcommand{\pt}{\partial t}
\newcommand{\pz}{\partial z}
\newcommand{\uvec}{\bs{u}}
\newcommand{\bvec}{\bs{B}}
\newcommand{\vort}{\bs{\omega}}
\newcommand{\F}{\bs{F}}
\newcommand{\T}{\tilde{T}}
\newcommand{\ez}{\bs{e}_z}
\newcommand{\ex}{\bs{e}_x}
\newcommand{\ey}{\bs{e}_y}
\newcommand{\eo}{\bs{e}_{\bs{\Omega}}}
\newcommand{\ppt}[1]{\frac{\partial #1}{\partial t}}
\newcommand{\ddt}[1]{\frac{d #1}{d t}}
\newcommand{\DDt}[1]{\frac{D #1}{D t}}
\newcommand{\ppts}[1]{\frac{\partial #1}{\partial t_s}}
\newcommand{\pptf}[1]{\frac{\partial #1}{\partial t_f}}
\newcommand{\ppz}[1]{\frac{\partial #1}{\partial z}}
\newcommand{\ddz}[1]{\frac{d #1}{d z}}
\newcommand{\ppzetas}[1]{\frac{\partial^2 #1}{\partial \zeta^2}}
\newcommand{\ppzs}[1]{\frac{\partial #1}{\partial z_s}}
\newcommand{\ppzf}[1]{\frac{\partial #1}{\partial z_f}}
\newcommand{\ppx}[1]{\frac{\partial #1}{\partial x}}
\newcommand{\ppxi}[1]{\frac{\partial #1}{\partial x_i}}
\newcommand{\ppxj}[1]{\frac{\partial #1}{\partial x_j}}
\newcommand{\ppy}[1]{\frac{\partial #1}{\partial y}}
\newcommand{\ppzeta}[1]{\frac{\partial #1}{\partial \zeta}}


\maketitle 

\setlength{\parindent}{0pt}
\setcounter{section}{1}

\section{Hydrodynamics Review}

What is a fluid?
\begin{itemize}
    \item Its flows! 
    \item it deforms continuously
\end{itemize}

Categorization of fluids:
\begin{itemize}
    \item Compressible v. incompressible
    \item viscous v. inviscid
    \item + many more
\end{itemize}

\vspace{20pt}

\bmp{.49}
    \centering
    {\Large \textbf{Eularian}}
    \vspace{5pt}

    Rate of change at a given point, no bother for where the fluid goes. 
    \begin{gather*}
        \frac{\partial}{\partial t}\left(\cdot\right) 
    \end{gather*}
\emp
\hspace{5pt}
\bmp{.49}
    \centering
    {\Large \textbf{Lagrangian}}
    \vspace{5pt}

    Follows the particle, introduces the advection term
    \begin{gather*}
        \uvec\cdot\grad \left(\cdot\right)
    \end{gather*}
\emp

\vspace{20pt}
{\Large \textbf{Mass Conservation}}
\vspace{5pt}


In order to conserve mass we consider an arbitrary eularian volume (i.e. the
volume doesn't move with the flow). We then find the total mass which is equal
to the integral of the density over the volume, and then consider the flux of
mass through the boundary (change in mass over time). Using the divergence
theorem, we then have a conservation equation for mass. 

\begin{gather*}
    \ppt{}\int_D\rho dV = \int_{\partial D} \rho\uvec \cdot \eta dA\\
    \ppt{\rho} + \grad\cdot \rho\uvec = 0 \\
    \ppt{\rho} + \uvec\cdot\grad\rho + \rho\grad\cdot\uvec = 0\\
    \frac{D\rho}{Dt} = - \rho \grad\cdot\uvec
\end{gather*}

More importantly, if we consider an incompressible fluid, i.e. $\rho = \rho_0$,
we have very specifically,
\begin{gather}
    \grad \cdot \uvec = 0
\end{gather}


\vspace{20pt}
{\Large \textbf{Stresses}}
\vspace{5pt}

Stresses can be divided into two catagories, body forces and surface forces.
Body forces are forces such as gravity and the electric force, which surface forces are
forces such as normal force and friction. 

\vspace{20pt}
{\Large \textbf{Newtons Second Law}}
\vspace{5pt}
Newton's second law 
\begin{gather*}
    \ppt{p} = \sum_i F_i
\end{gather*}
where $p$ here is the momentum of a fluid parcel. In actuality, the momentum can
be written as $p = \int_D \rho\uvec dV$. So, 
\begin{gather*}
    \DDt{}\int_D \rho\uvec dV = \int_D \rho \bs{g} dV + (\text{other body
    forces}) + \grad \cdot \bs{\tau}\\
    \rho\DDt{\uvec} = \rho F + \grad\cdot\bs{\tau}
\end{gather*}
where $\bs{\tau}$ is the stress tensor acting on the fluid parcel. Surface
forces are then introduced into this stress tensor. First and foremost, surface
pressure is introduced along the stress tensor. 
\begin{gather*}
    \tau_{ij} = -p\delta_{ij} + \sigma_{ij}
\end{gather*}
where $\sigma_{ij}$ is the deviatoric stress tensor and is responsible for the
off-diagonal components of the stress tensor. Some components are the velocity
gradient tensors, $\ppxj{u_i}$ and $\ppxi{u_j}$. Each of these has a symmetric
compoenent and an antisymmetric componenet. 
\begin{gather*}
    \ppxj{u_i} = \frac{1}{2}\left(\ppxj{u_i} + \ppxi{u_j}\right) +
    \frac{1}{2}\left(\ppxj{u_i} - \ppxi{u_j}\right)
\end{gather*}
The first term is labeled symmetric and denoted $e_{ij}$ while the second
component is the rotation componenet. 



\section{Continuing Review of the Kinematic Equations}

\subsection{Obtaining the Navier-Stokes equation}

Decomposition of the deviatoric stress tensor reveals a 4th order tensor with 81
components.
\begin{gather*}
    \bs{\sigma}_{ij} = \bs{A}_{ijkl}\bs{e}_{kl}
\end{gather*}
In order to reduce the complexity of the system, we make some assumptions about
the tensor $\bs{A}_{ijkl}$. First, we state that this tensor must be isotropic,
i.e. that it doesn't care about the direction of the stress with respect to the
coordinate system it is in. We have, 
\begin{gather*}
    \bs{A}_{ijkl} = \mu \bs{\delta}_{ij}\bs{\delta}_{kl} +
    \mu'\bs{\delta}_{ik}\bs{\delta}_{jl} +
    \mu''\bs{\delta}_{il}\bs{\delta}_{jk}
\end{gather*}
Next, we assume that this tensor must be symmetric. This reduces the complexity
down to two coefficients, $\mu$, the viscosity, and $\mu'$ which is the bulk viscosity. 

In order to obtain the Navier-Stokes equation, we require the Stokes assumption
which postulates that the diagonal components of the deviatoric stress tensor
are zero, i.e. $\bs{\sigma}_{ii} = 0$. 
\begin{gather*}
    \bs{\sigma}_{ij} = 2\mu\left(\bs{e}_{ij} -
    \frac{1}{3}\left(\grad\cdot\uvec\right)\bs{\delta}_{ij}\right)    \\
    \bs{\tau}_{ij} = -p\bs{\delta}_{ij} + 2\mu\left(\bs{e}_{ij} -
    \frac{1}{3}\left(\grad \cdot \uvec\right)\bs{\delta}_{ij}\right)
\end{gather*}
Therefore, when we take the divergence of this stress tensor we obtain the
Navier-Stokes equation:
\begin{gather*}
    \rho\DDt{\uvec} = \rho\bs{F} - \grad p + \mu\left[ \grad^2 \uvec +
    \frac{1}{3}\grad \left(\grad \cdot \uvec\right)\right]
\end{gather*}
Of course, when working in an incompressible framework (i.e. $\grad\cdot\uvec =
0$), we have that part of the diffusive term disappears from the equation,
resulting in the commonly used equation:
\begin{gather}
    \label{eq:NS}
    \DDt{\uvec} = \bs{F} - \frac{1}{\rho_0}\grad p + \nu\grad^2 \uvec
\end{gather}
Additional terms are included in this equation as necessary to model relevant
physics of various fluid systems. For example, if in a rotating frame we include
the coriolis force $2\Omega(\bs{e}_{\Omega} \times \uvec)$, if some component of
the fluid is stratified we need some buoyancy forcing $T/N^2\bs{e}_z$. And most
relevant, there might be magnetic forces which affect the fluid, in which case
we obtain the MHD equations. 

\subsection{Vorticity equation}

Vorticity is a quantity related to the fluid field which can be very important
to the scientific study of fluid dynamics. The vorticity is obtained by taking
the curl of the velocity field. 
\begin{gather*}
    \bs{\omega} = \grad \times \uvec
\end{gather*}
The vorticity has an evolution-advection equation just as the velocity field
does, and in fact the vorticity equation is obtained by taking the curl of the
Navier-Stokes equations.

\begin{gather*}
    \grad\times\eqref{eq:NS}\\
    \DDt{\vort} =  \left(\vort\cdot\grad\right)\uvec +
    \vort\left(\grad\cdot\uvec\right) + \grad\times\F -
    \frac{1}{\rho^2}\grad\rho\times\grad p + 
    \nu\grad^2\vort 
\end{gather*}

If the flow is incompressible, one of the vortex stretching terms disappears.
Generally, the first two right hand terms are vortex stretching/tilting/speed-up
terms. Then the pressure and density gradient cross product is the baroclinicity
term, the curl of $\F$ is the forcing of vorticity, and finally, we have a
viscous diffusion of vorticity which behaves similarly to the diffusion of
velocity. 


Baroclinicity is perhaps the most unintuitive term in this equation, and it
simply represents the creation of rotation in the fluid when there is a
disalignment between the pressure and density gradients in the fluid. 

Some fluid dynamicists prefer to study the vorticity equation, especially for
rotating flows where voricites and cyclones are common phenomenon in the flow
field. 

\subsection{Rotation}

In the presence of rotation, the coriolis force becomes relevant as the motion
of a fluid particle is deflected due to the rotation of the cordinate frame.
That is, our equations are modified such that, 
\begin{gather*}
    \ppt{\bs{q}}_F = \ppt{\bs{q}}_R + 2\bs{\Omega}\times \bs{q} -
    \bs{\Omega}^2\bs{R}
\end{gather*}

This also introduces an additional term to the vorticity equation which looks
like, $+\left(2\bs{\Omega}\cdot\grad\right)\uvec$. 


\section{Conservation of Energy and Maxwell's equations}

\subsection{Conservation of Energy}

The equation of state chosen for a particular problem is a source of physics
which affects the solutions of a given PDE. The incompressible equation of
state is used very commonly as an equation of state. Another common one is the
ideal gas law $pV = \rho RT$. 

In order to understand the origin and importance of the equation of state, the
laws of thermodynamics are needed. 

The first law of thermodynamics states, 

\begin{gather*}
    \ppt{e} = \ppt{W} + \ppt{Q}
\end{gather*}
where $e$ is the internal energy, $W$ is work done on the system, and $Q$ is
heat flux into the system. However, for a fluid flow taken from a Lagrangian
perspective, we must modify this law of thermodynamics. It must include the
energy given by the velocity field. 

\begin{gather*}
    \DDt{}\int_D \rho\left(e+\frac{1}{2}\uvec^2\right)dV = \int_D \rho
    \F\cdot\uvec dV + \int_{\partial D} \bs{\tau}\cdot\uvec dS - \int_{\partial
    D} \bs{q}\cdot dS \\
    \rho\DDt{}\left(e+\frac{1}{2}\uvec^2\right) = \rho\F\cdot\uvec +
    \grad(\bs{\tau}\cdot\uvec) -
    \grad\cdot\bs{q} 
\end{gather*}

Next, we obtain a mechanical energy equation by dotting $\uvec$ by the
Navier-Stokes equation and adding $\uvec^2/2 \cdot \DDt{\rho}$
\begin{gather*}
    \DDt{\rho\uvec^2/2} = \rho\F\cdot\uvec -
    \uvec\cdot\left(\grad\cdot\bs{\tau}\right) + \ldots\\
    \uvec\cdot\left(\grad\cdot\bs{\tau}\right) = 
    \Phi = 2\mu\left[\bs{e}_{ij} -
    \frac{1}{3}(\grad\cdot\uvec)\bs{\delta}_{ij}\right]
\end{gather*}

Finally, we obtain an energy equation with a positive definite dissipation term
$\Phi$ which acts purely to remove energy from the system. 
\begin{gather*}
    \rho\DDt{e} = -\grad \cdot \bs{q} - p\left(\grad\cdot\uvec\right) + \Phi
\end{gather*}

The Second law of Thermodynamics also plays an important role in the
conservation of energy. The ssecond law makes statements about the entropy of a
system, $S$.
\begin{gather*}
    dS = \frac{dq}{T}\\
    TdS =  de + pdV\\
    T\ddt{S} = \ddt{e} - \frac{p}{\rho^2}\ddt{\rho}\\
    \rho \DDt{S} = - \frac{\grad \cdot \bs{q}}{T} + \frac{k}{T^2}|
    \grad T|^2 + \mu\frac{\Phi}{T}
\end{gather*}
Essentially, since both $|\grad T|^2$ and $\Phi$ are positive definite terms and
their coefficients are positive definite, it must be that the entropy of a
system can only increase ``on average'' (curse the statisticians). 

\subsection{Maxwell's Equations and MHD}

Electricity and Magnetism are very closely related to one another and governed
by a main set of governing equations. The main variables which we consider are a
position vector, $\bs{x}$, a velocity field, $\uvec$, density $\rho$, pressure
$p$, time $t$, temperature $T$, magnetic field (magnetic flux density) $\bvec$,
magnetic field strength $\bs{H}$, electric field $\bs{E}$, electric displacement
$\bs{D}$, electric current density $\bs{j}$, and charge density $\rho_e$. 

Alongside these variables, we have constants describing components of
electropmangetism: permitivity $\varepsilon$, permeability $\mu$, and
conductivity $\sigma$. Permitivity describes the charge requirement for a
specific electric field, i.e. large $\varepsilon$ implies a larger charge is
needed for a specific electric field. Permeability describes the current
requirement for a specific magnetic field, i.e. large $\mu$ implies a smaller
current is needed to obtain a specific magnetic field. 

Consitutive relationships describe the relationships between specific
electromagnetic quantities. 
\begin{gather*}
    \bs{H} = \frac{\bvec}{\mu}, \text{ for an isotropic permeability}\\
    \bs{D} = \varepsilon \bs{E}, \text{ for an isotropic permitivity}
\end{gather*}
where generally, we take $\mu = \mu_0$ and $\varepsilon = \varepsilon_0$ where
$q_0$ is taken from a vacuum. 

Now we write Maxwell's equations in their differential form:
\begin{align*}
    &\grad\cdot\bvec = 0, \text{ Gauss' law for magnetism}
    & \grad \cdot \bs{E} = \frac{\rho_e}{\varepsilon_0}, \text{ Gauss' law}\\
    &\grad \times \bs{E} = -\ppt{\bvec}, \text{ Faraday's law}
    &\grad \times \bvec = \mu_0\left(\bs{j} + \varepsilon_0\ppt{\bs{E}}\right),
    \text{ Ampere's Law of induction}
\end{align*}

They can be written in their integral form as well: 
\begin{align*}
    &\oint_{\partial D} \bvec\cdot\bs{dA} = 0 
    & \oint_{\partial D} \bs{E}\cdot\bs{dA} = \frac{q}{\varepsilon_0}\\
    &\oint_L \bs{E}\cdot\bs{dL} = -\ppt{\phi_{B}} 
    & \oint_L\bvec\cdot\bs{dL} = \mu_0\bs{j} + \mu_0\varepsilon_0\ppt{\phi_E}
\end{align*}
where $\phi_B = \int_S \bvec\cdot\bs{dA}$ is the total magnetic fluid, and 
$\phi_E = \int_S \bs{E}\cdot\bs{dA}$ is the total electric flux. 


We cover their derivations in a brief sense also. Consider a positive point
charge which creates an electric field. This imposes a force acting on any other
point charge in the field. This force is called Coulomb force given by $F =
q_1q_2/(4\pi\varepsilon r^2)$. Thus we have a given electric field of strength
$E/q = q_1/(4\pi\varepsilon r^2)$. We obtain the total electric flux $\phi_E$ 

\begin{gather*}
    d\phi_E = \bs{E}\cdot\bs{dA}\\
    \phi_E = \oint_S\frac{q_1}{4\pi\varepsilon r^2}\cdot\bs{dA}\\
    \phi_E = \frac{q_1}{4\pi\varepsilon r^2}\oint_S\bs{dA}\\
    \phi_E = \frac{q_e}{\varepsilon}
\end{gather*} 
where $q_e$ in the final equation is given by the sum of all point charges
enclosed in the closed volume, i.e. $q_e = \sum_i q_i$. Notice that $q_e$ can be
thought of as mass for point charges, i.e. the integral of the charge density
equals the total charge similar to how the integral of mass density equals the
total mass. It can be represented as a sum of point charges because point
cahrges are discrete and do not usually exist in a continuum. 

Finally, using the divergence theorem, 
\begin{gather*}
    \oint_S \bs{E}\cdot\bs{dA} = \int_V \grad\cdot\bs{E} dV\\
    q_e = \int_V \rho_{E} dV\\
    \grad\cdot\bs{E} = \frac{\rho_E}{\varepsilon}
\end{gather*}
Similarly, the same proof holds for Gauss' law of magnetism, only that monopoles
do not exist in magnetic fields, i.e. every source must have a sink. Therefore,
for an arbitraty volume it must be that the divergence of the magnetic field
must be zero: 
\begin{gather*}
    \div\bvec = 0
\end{gather*}
\end{document}
